---
title: "Names and Values"
format: html
editor: visual
---

```{r}
library(lobstr)
```

## Binding Basics

```{r}
x <- c(1, 2, 3)
```

R does two things in the above line - creating an object, a vector of values - binding that object to a name, `x`

The object, or value, doesn't have a name, it's the name that has a value. So you can think of a name as a reference for a value.

So when we run

```{r}
y <- x
```

we don't get a new copy of the object, but a second binding to that original object, we can verify this by looking at the object address.

```{r}
obj_addr(x) == obj_addr(y)
```

### Exercises

#### 1

Explain the relationship between a, b, c and d in the following code:

```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
```

a, b, and c are all references to the same object, but d is a reference to an object with the same values but with a different address.

#### 2

The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? 

```{r}
mean
base::mean
get("mean")
evalq(mean)
match.fun("mean")

obj_addr(mean)
obj_addr(base::mean)
obj_addr(get("mean"))
obj_addr(evalq(mean))
obj_addr(match.fun("mean"))
```


They all do point to the same underlying function object


#### 3

By default, base R data import functions, like `read.csv()`, will automatically convert non-syntactic names to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?

This might be problematic when the column names are actually data points, this name conversion thus corrupts the data. You can use `check.names = F`


## Copy on Modify

```{r}
x <- c(1, 2, 3)
y <- x

y[[3]] <- 4
```


Here x and y originally point to the same object, but when y is modified R makes a copy of the original object, this time with one item changed and rebinds y to that object. It can help to think of R objects as immutable or unchangeable.

Like vectors, lists use copy-on-modify behavior; the original list is left unchanged, and R creates a modified copy. This, however, is a shallow copy: the list object and its bindings are copied, but the values pointed to by the bindings are not.

Data frames are lists of vectors, so copy-on-modify has important consequences when you modify a data frame.

```{r}
d1 <- data.frame(x = c(1, 5, 6), y = c(2, 4, 3))
d2 <- d1
d2[, 2] <- d2[, 2] * 2
```

since we only modified one column then only that column must be modified, all other bindings point to the original objects

### Exercises 

#### 1

Why is tracemem(1:10) not useful?

It isn't useful because there is no name that points to that object. When 1:10 is called an object with an address is indeed created, but since it has no name that points to it, it isn't useful to track the object.

#### 2

Explain why tracemem() shows two copies when you run this code. Hint: carefully look at the difference between this code and the code shown earlier in the section.

```{r}
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4
```

We see copy on modify behavior here because the original values in the object are integers and the modification is to change one of those values to a double.

#### 3

Sketch out the relationship between the following objects:

```{r}
a <- 1:10
b <- list(a, a)
c <- list(b, a, 1:10)
```

a references the object 1:10 with an address say j. b references an object where each item in the object references object j. finally c has a reference to b and a reference to a as well as a reference to a new object 1:10 with a different address.


## Object Size

```{r}
x <- runif(1e6)
obj_size(x)

y <- list(x, x, x)
obj_size(y)
```

note that these are almost the same size because of the fact that elements in lists are references to values.

## Modify in place

If an object has a single name bound to it, R will modify it in place:

```{r}
v <- c(1, 2, 3)
v[[3]] <- 4
```

after the modify, `v` continues to bind to the same object. It can be hard to predict whether or not a copy will occur.

We can see this with an example with for loops: 

```{r}
x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
medians <- vapply(x, median, numeric(1))

for (i in seq_along(medians)) {
  x[[i]] <- x[[i]] - medians[[i]]
}

cat(tracemem(x), "\n")
#> <0x7f80c429e020> 

for (i in 1:5) {
  x[[i]] <- x[[i]] - medians[[i]]
}
```

We can see that with each iteration of the for loop, the data frame is copied 3 times.

While it’s not hard to determine when a copy is made, it is hard to prevent it. If you find yourself resorting to exotic tricks to avoid copies, it may be time to rewrite your function in C++

### Environments

Environments are always modified in place!

## Unbinding and Garbage collector

Here we create two objects, but once this chunk is run, neither object is bound to a name

```{r}
x <- 1:3
x <- 2:4

rm(x)
```

How do these objects get deleted? That’s the job of the garbage collector, or GC for short. The GC frees up memory by deleting R objects that are no longer used, and by requesting more memory from the operating system if needed.